---
title: "Predictive Assignment"
author: "Davyd"
date: "21 May 2018"
output: html_document
---

####Background 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

####Data Source

Training data can be downloaded from (https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

Testing data can be downloaded from (https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)


####Load Required Packages

```{r, cache = T}
library(caret)
library(lattice)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

### Download the Data

```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```  

####Load the data

```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```

####Data Cleaning

This step removes the unnecessary variables and NAs

```{r, cache = T}
sum(complete.cases(trainRaw))
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```

####Cross Validation

This step divides the training data into a training set and cross validation set via 70/30 proportion

```{r, cache = T}
set.seed(12345) 
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

####Data Modelling

This step uses a Random Forest model with 5-fold cross validation

```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```

Next, we estimate the performance of the model on the validation dataset.  

```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```
```{r, cache = T}
accuracy <- postResample(predictRf, testData$classe)
accuracy
oose <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
oose
```

#### Predicting for Test Dataset

This step applies the model to the test dataset

```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```  


####Appendix

1. Correlation Matrix Plot 

```{r, cache = T}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```

2. Decision Tree 

```{r, cache = T}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) 
```